---
title: "Nitrosomonas ureae fast-track system standardisation"
author: "Alex Kanellopoulos"
date: "2024-11-26"
output: html_document
---

Brief description: This is a script that includes all the necessary code to reproduce the results from the respective manuscript. 

Step 1: Let's see all the necessary packages needed for the analysis and load them. If they are not installed, we will install them first. 

```{r Library Installation and Loading, echo=TRUE, message=FALSE, warning=FALSE}
# Here we store the necessary packages in the necessary_packages variable
necessary_packages <- c("readxl","sicegar","ggplot2","viridis","dplyr","SciViews","rstatix","agricolae","tibble","patchwork","scales")

# Here we check if the necessary packages are installed. If not, R installs them 
if (!requireNamespace(necessary_packages, quietly = TRUE)) {
  install.packages(necessary_packages)
}

# Here, for every package included in the necessary_packages variable, R will load it. The character.only = TRUE ensures that library() will try to interpret the package names as characters (as it needs for it to work) 
for (package in necessary_packages) {
  library(package, character.only = TRUE)
}
```

**Section 1 : N. ureae routine culture growth curve**

The μmax can be found in Supplementary Table S5.The μmax calculation is shown in Supplementary Figure S4C. The routine culture growth model curve is used in Supplementary Figure S2C  

Step 1: Data input.

```{r Data import and pre-processing, echo=TRUE}
# Here, we create an R data frame named with the data
library(readxl)
data_routine <- read_excel("NUE Standardisation.xlsx", sheet = 1, col_names = TRUE)
# We can change the names of the columns to more R-friendly names
names(data_routine) <- c("Time", "Replicate", "Generation", "Nitrite", "Mean", "St.Dev", "CV%")
#We keep only the rows that contain a Nitrite value (thus removing rows that are left blank to make life easier when viewing the .xlsx file)
data_routine <- filter(data_routine, !is.na(Nitrite))
#Let's view the data
glimpse(data_routine)
```

Step 2: We can use a single - sigmoidal model and fit our data in this in order to obtain an equation describing the culture growth (in this case synonymous with nitrite production) over time

```{r echo=TRUE, message=FALSE, warning=FALSE}
# We want to keep only the 2 columns containing the Time and Nitrite values
data_model <- data.frame(data_routine$Time, data_routine$Nitrite)
# We change the names of the two columns to names required by the "sicegar" package
names(data_model) <- c("time", "intensity")
glimpse(data_model)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Let's give the command to generate the model
library(sicegar)
fitObj_sm <- fitAndCategorize(dataInput = data_model,
                              threshold_t0_max_int = 100, 
                              threshold_minimum_for_intensity_maximum = 0.02)
```

Step 3: We can extract the desired model parameters

```{r echo=TRUE, message=FALSE, warning=FALSE}
I_max <- fitObj_sm$sigmoidalModel$maximum_Estimate
a <- fitObj_sm$sigmoidalModel$slopeParam_Estimate
t_mid <- fitObj_sm$sigmoidalModel$midPoint_Estimate

```

Step 4: We can overlay the routine cultures growth curves with the model. This will help identify the exponential phase to calculate the maximum specific growth rate (μmax) 

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Prepare the data
data_routine_line_plot <- filter(data_routine, !is.na(Mean))
# Prepare the sigmoidal model equation using the model parameters
model_curve <- function(t) {I_max/(1+exp(-a*(t-t_mid)))}
# Prepare the legend order and short the generations 
legend_order <- c("1st", "2nd", "3rd", "4th", "5th", "6th")
data_routine_line_plot$Generation <- factor(data_routine_line_plot$Generation, levels = legend_order)
# We will use an activity plot but overlay the model curve
library(ggplot2)
library(viridis)
activity_line_plot_with_model <- ggplot(data_routine_line_plot, aes(x = Time, y = Mean)) + 
  geom_point(aes(colour = Generation)) + 
  theme_light() + 
  labs(x = "Time (Hours)", y = "[Nitrite] (μM)") + 
  theme(axis.title = element_text(size = 28), axis.text = element_text(size = 24), legend.text = element_text(size = 20), legend.title = element_text(size = 20)) + 
  scale_y_continuous(breaks = seq(0,max(data_routine_line_plot$Mean), by = max(data_routine_line_plot$Mean)/5)) + ylim(0,max(data_routine_line_plot$Mean)) + 
  scale_color_viridis(option = "D", discrete = T)

activity_line_plot_with_model <- activity_line_plot_with_model + 
  geom_errorbar(aes(ymin = Mean - St.Dev, ymax = Mean + St.Dev, width = 0.2, colour = Generation)) +
  geom_line(data = data_routine_line_plot, aes(x = Time, y = Mean, colour = Generation), linewidth = 0.6) +
  stat_function(fun = model_curve, linewidth = 2, aes(colour = "Growth Model Curve"))

print(activity_line_plot_with_model)
```

Step 5: We can calculate the maximum specific growth rate from the activity data. Briefly, we isolate the data from the exponential growth phase (20h - 150h) and we perform a semi - logarithmic transformation, using the natural logarithm (ln) to transform the nitrite values 

```{r echo=TRUE, message=FALSE, warning=FALSE}
# We assume that the exponential growth phase lasts from 20 h - 150 h, so we isolate these data
library(dplyr)
data_exponential <- filter(data_routine, between(Time, 20, 150))

# Prepare the dataframe 
data_exponential$Generation <- factor(data_exponential$Generation, levels = legend_order)


# We transform the Nitrite values 
library(SciViews)
data_exponential_ln <- mutate(data_exponential, Nitrite = ln(Nitrite))

# We will evaluate the quality of the linear regression
# The maximum specific growth rate is congruent with the slope coefficient of the linear regression
linear_regression_report <- summary(lm(Nitrite ~ Time, data = data_exponential_ln))
μmax <- round(linear_regression_report$coefficients[2,1],3)
μmax_std_error <- round(linear_regression_report$coefficients[2,2],3)
linear_regression_report
μmax
μmax_std_error

# We plot the data and we will incorporate the maximum specific growth rate value in the plot
line_plot_exponential_ln <- ggplot(data_exponential_ln, aes(x = Time, y = Nitrite)) +
  theme_light() + 
  geom_point(aes(colour = Generation)) + 
  labs(y = "ln[Nitrite]", x = "Time (Hours)") + 
  theme(axis.title = element_text(size = 28), axis.text = element_text(size = 24), legend.text = element_text(size = 20), legend.title = element_text(size = 20)) + 
  scale_color_viridis(option = "D", discrete = T) +
  stat_smooth(method = "lm", col = "blue", se = TRUE, linewidth = 1, linetype = "dashed") + 
  geom_text(x = 90, y = 5, label = paste("μmax =", μmax, "±", μmax_std_error), colour = "blue", size = 4.5, hjust = 0) + 
  geom_text(x = 90, y = 4.5, label = paste("R-squared = ", round(linear_regression_report$adj.r.squared,2)), colour = "blue", size = 4.5, hjust = 0)

print(line_plot_exponential_ln)

ggsave("Figure S4C-NUE maximum specific growth rate.pdf", plot = line_plot_exponential_ln, device = "pdf", height = 10, width = 10)

```



**Section 2 : N. ureae fast-track cultures cell density data**

The boxplots produced can be found in Supplementary Figure S2A and Supplementary Figure S2B. Cell density data are found in Supplementary Table S4 

Step 1: Data input. 

```{r echo=TRUE, message=FALSE, warning=FALSE}
# We inport the data
library(dplyr)
data_1D_DNA <- read_excel("NUE Standardisation.xlsx", sheet = 2, col_names = TRUE)
names(data_1D_DNA) <- c("Step", "Replicate","Repetition", "Cells", "Mean", "St.Dev")
glimpse(data_1D_DNA)

# we create the legend order with the different steps of the harvesting procedure and short the Step column of the dataframe
legend_order_DNA <- c("Routine Culture Initial", "Routine Culture Final", "Inocula after cell harvesting", "Merged Inoculum", "Concentrated Culture Initial")

data_1D_DNA$Step <- factor(data_1D_DNA$Step, levels = legend_order_DNA)

```


Step 2: Statistical analysis

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(rstatix)
library(agricolae)
library(tibble)

# Shapiro distribution test 
shapiro_DNA <- shapiro_test(data_1D_DNA, Cells)
shapiro_DNA

# Kruskal-Wallis test
kruskal_DNA <- with(data = data_1D_DNA, kruskal(Cells, Step, p.adj = "bonferroni", group = T, alpha = 0.05))
print(kruskal_DNA$groups)

# Short the Kruskal-Wallis results by the legend order created for the steps of the harvesting procedure so the groups will be assigned correctly to the respective boxplot
kruskal_results <- data.frame(kruskal_DNA$groups)
kruskal_results <- rownames_to_column(kruskal_results, var = "Step")
treatment_order <- legend_order_DNA
kruskal_results$Step <- factor(kruskal_results$Step, levels = treatment_order)
kruskal_results <- kruskal_results[order(kruskal_results$Step), ]
View(kruskal_results)

data_1D_DNA_means <- summarise(group_by(data_1D_DNA, Step), Mean = mean(Cells))
data_1D_DNA_means$Step <- factor(data_1D_DNA_means$Step, levels = treatment_order)
data_1D_DNA_means <- data_1D_DNA_means[order(data_1D_DNA_means$Step), ]

View(data_1D_DNA_means)

```


Step 3: Prepare the boxplot (Figure S2A)

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(ggplot2)
box_plot_DNA <- ggplot(data_1D_DNA, aes(x = Step, y = Cells)) + 
  geom_boxplot(aes(colour = Step), width = 5) + theme_light() + 
  labs(title = "N. ureae Cell Density Determination", x = "Sample", y = "Cell Density (cells/mL)") +
  theme(plot.title = element_text(size = 20), axis.title = element_text(size = 28), axis.text = element_text(size = 24), legend.text = element_text(size = 20), legend.title = element_text(size = 20), plot.subtitle = element_text(size = 18), axis.text.x=element_text(angle = 45, hjust = 1)) + 
  scale_x_discrete(limits = legend_order_DNA) +
  scale_y_continuous(breaks = seq(1e+07, 1.9e+09, by = 2e+08)) + 
  scale_color_manual(values = c("Routine Culture Initial" = "#006b6b", "Routine Culture Final" = "#00ffff", "Inocula after cell harvesting" = "#ff6bff", "Merged Inoculum" = "#b200b2", "Concentrated Culture Initial" = "#840084"), breaks = legend_order_DNA, labels = legend_order_DNA) + 
  geom_hline(yintercept = 1.0e+07, linetype = "dashed") + 
  geom_text(data = data.frame(kruskal_results), aes(label = groups, y = data_1D_DNA_means$Mean), vjust = -0, size = 10)

print(box_plot_DNA)
```


Step 4: Prepare the lineplot (Figure S2B)

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(ggplot2)
data_1D_DNA_lineplot <- filter(data_1D_DNA, !is.na(data_1D_DNA$Mean))

line_plots_DNA <- ggplot(data_1D_DNA_lineplot, aes(x = factor(Step), y = Mean, group = Repetition)) + geom_point(aes(colour = Repetition)) + 
  theme_light() + 
  labs(title = "Nitromonas ureae Cell Density Determination",subtitle = "Comparison between consecutive attempts", x = "Sample", y = "Cell Density (cells/mL)") + 
  theme(axis.title = element_text(size = 40), axis.text = element_text(size = 36), legend.text = element_text(size = 32), legend.title = element_text(size = 36), axis.text.x=element_text(angle = 45, hjust = 1), plot.title = element_text(size = 20), plot.subtitle = element_text(size = 18), legend.position = "right") + 
  scale_x_discrete(limits = legend_order_DNA) + 
  theme(axis.text.x=element_text(angle = 45, hjust = 1)) + 
  scale_y_continuous(breaks = seq(1e+07, 1.9e+09, by = 2e+08)) + 
  geom_hline(yintercept = 1.0e+07, linetype = "dashed") + 
  scale_color_manual(values = c("1st" = "#840084", "2nd" = "#b200b2", "3rd" = "#ff6bff", "4th"= "#006b6b"))

line_plots_DNA <- line_plots_DNA + 
  geom_line(data = data_1D_DNA_lineplot, aes(colour = Repetition), linewidth = 1.6) +
  geom_errorbar(aes(ymin = Mean - St.Dev, ymax = Mean + St.Dev, colour = Repetition), width = 0.6)

print(line_plots_DNA)
```


**Section 3 : N. ureae fast-track cultures cell activity curves**

The activity rate can be found in Supplementary Table S5 and its calculation is shown in Supplementary Figure 4D. The fast-track culture activity plot is used in Supplementary Figure S2C  

Step 1: Data input.

```{r Data import and pre-processing, echo=TRUE}
# Here, we create an R data frame named with the data
library(readxl)
data_1D_activity <- read_excel("NUE Standardisation.xlsx", sheet = 3, col_names = TRUE)
# We can change the names of the columns to more R-friendly names
names(data_1D_activity) <- c("Time", "Replicate", "Repetition", "Nitrite", "Mean", "St.Dev", "CV%")
#Let's view the data
glimpse(data_1D_activity)
```

Step 2: Merge the fast-track cultures activity curves with the routine culture growth model.

```{r Data import and pre-processing, echo=TRUE}
# Prepare the legend order
legend_order_activity <- c("1st","2nd","3rd")

line_plot_merged <- ggplot(data_1D_activity, aes(x = Time, y = Nitrite)) + 
  geom_point(aes(colour = Repetition), size = 2) + 
  geom_smooth(aes(colour = Repetition), method = lm, linetype = "dashed", linewidth = 2, se = F, fullrange = T) + theme_light() + 
  labs(title = "Nitrosomonas ureae  fast - track culture", subtitle = "Linear Regression compared to the growth model for the Routine cultures", x = "Time (Hours)", y = "[Nitrite] (uM)") + 
  stat_function(fun = model_curve, aes(color = "Growth Model Routine Culture"), linewidth = 2) + 
  xlim(0,200) + theme(legend.position = "bottom", legend.box = "vertical", axis.title = element_text(size = 28), axis.text = element_text(size = 24), legend.text = element_text(size = 24), legend.title = element_text(size = 24), plot.title = element_text(size = 24), plot.subtitle = element_text(size = 22)) + 
  scale_color_manual(values = c("1st" = "#840084", "2nd" = "#b200b2", "3rd" = "#ff6bff", "Growth Model Routine Culture" = "#006b6b"), name = "Culture") + 
  theme(axis.title = element_text(size = 28), axis.text = element_text(size = 24)) +
  ylim(0,1000)

line_plot_merged <- line_plot_merged + 
  geom_text(x = Inf, y = -Inf, label = paste("R - squared =", round(summary(lm(Nitrite~Time, data_1D_activity))$adj.r.squared,4)), colour = "magenta4", size = 10, hjust = 1.1, vjust = -2) 

print(line_plot_merged)
```

Step 3: Print the whole Figure (Supplementary Figure S2)

```{r Data import and pre-processing, echo=TRUE}
# Prepare the patchwork
library(patchwork)
patchwork <- (box_plot_DNA + line_plots_DNA) / line_plot_merged + plot_annotation(tag_levels = "A") & 
  theme(plot.tag = element_text(size = 32)) 
print(patchwork)

cairo_pdf("Figure S2. NUE_Standardisation.pdf", height = 30, width = 30)
print(patchwork)
dev.off()
```

Step 4: Determine the activity rate of the fast-track cultures in a similar way to the routine cultures μmax (see Section 1 - Step 5 for guides)

```{r Data import and pre-processing, echo=TRUE}

library(SciViews)
data_1D_ln <- mutate(data_1D_activity, Nitrite = ln(Nitrite))

linear_regression_report <- summary(lm(Nitrite ~ Time, data = filter(data_1D_ln, !Nitrite == -Inf)))
μmax <- round(linear_regression_report$coefficients[2,1],3)
μmax_std_error <- round(linear_regression_report$coefficients[2,2],3)
linear_regression_report
μmax
μmax_std_error

library(scales)
library(viridis)
line_plot_exponential_ln <- ggplot(data_1D_ln, aes(x = Time, y = Nitrite)) +
  theme_light() + 
  geom_point(aes(colour = Repetition)) + 
  labs(y = "ln[Nitrite]", x = "Time (Hours)") + 
  theme(axis.title = element_text(size = 28), axis.text = element_text(size = 24), legend.text = element_text(size = 20), legend.title = element_text(size = 20)) + 
  scale_color_viridis(option = "D", discrete = T) +
  stat_smooth(method = "lm", col = "blue", se = TRUE, linewidth = 1, linetype = "dashed") + 
  geom_text(x = 10, y = 4, label = paste("Activity rate =", μmax, "±", μmax_std_error), colour = "blue", size = 6, hjust = 0) + 
  geom_text(x = 10, y = 3.70, label = paste("R-squared = ", round(linear_regression_report$adj.r.squared,2)), colour = "blue", size = 6, hjust = 0)

print(line_plot_exponential_ln)

ggsave("Figure S4D-NUE fast-track activity rate.pdf", plot = line_plot_exponential_ln, device = "pdf", height = 10, width = 10)

```

This is the end of this script. 
Please always remember to cite the authors of any code you use!
