---
title: "Nitrososphaera viennensis fast-track system standardisation"
author: "Alex Kanellopoulos"
date: "2024-11-27"
output: html_document
---

Brief description: This is a script that includes all the necessary code to reproduce the results from the respective manuscript. 

Step 1: Let's see all the necessary packages needed for the analysis and load them. If they are not installed, we will install them first. 

```{r Library Installation and Loading, echo=TRUE, message=FALSE, warning=FALSE}
# Here we store the necessary packages in the necessary_packages variable
necessary_packages <- c("readxl","viridis","scales","dplyr","ggplot2","agricolae","sicegar","patchwork","gridExtra","rstatix", "tibble")

# Here we check if the necessary packages are installed. If not, R installs them 
if (!requireNamespace(necessary_packages, quietly = TRUE)) {
  install.packages(necessary_packages)
}

# Here, for every package included in the necessary_packages variable, R will load it. The character.only = TRUE ensures that library() will try to interpret the package names as characters (as it needs for it to work) 
for (package in necessary_packages) {
  library(package, character.only = TRUE)
}
```

**Section 1 : "N. viennensis routine culture growth curve and amoA abundance curve**

The routine culture growth model curve is used in Supplementary Figure S6Β 

Step 1: Data input.

```{r Data import and pre-processing, echo=TRUE}
# Here, we create an R data frame named with the data
library(readxl)
data_routine <- read_excel("NVE Standardisation.xlsx", sheet = 1, col_names = TRUE)
# We can change the names of the columns to more R-friendly names
names(data_routine) <- c("Time", "Nitrite", "Mean_Nitrite", "St.Dev_Nitrite", "Cells", "Mean_Cells","St.Dev_Cells")
#We keep only the rows that contain a Nitrite value (thus removing rows that are left blank to make life easier when viewing the .xlsx file)
data_routine <- filter(data_routine, !is.na(Nitrite))
#Let's view the data
glimpse(data_routine)
```

Step 2: We can use a double curve plot to check congruence between nitrite concentration and amoA abundance

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Prepare the plot in steps and export
cairo_pdf("Figure S6B-NVE_nitrite_amoA_congruence.pdf", width = 10, height = 8)

# Set plot parameters
par(mar=c(5, 4, 4, 6) + 0.1)

plot(filter(data_routine, !is.na(Mean_Nitrite))$Time, filter(data_routine, !is.na(Mean_Nitrite))$Mean_Nitrite, pch=16, axes=FALSE, ylim=c(0,2100), xlab="", ylab="", 
     type="b",col="black", main="Nitrososphaera viennensis - Culture activity compared to flow cytometry data")
axis(2, ylim=c(0,2100),col="black",las=1)  
mtext("[Nitrite] (μM)",side=2,line=3)
box()

par(new=TRUE)

plot(filter(data_routine, !is.na(Mean_Nitrite))$Time, filter(data_routine, !is.na(Mean_Nitrite))$Mean_Cells,  xlab="", ylab="", ylim=c(4E+04,3.5E+07), 
     axes=FALSE, type="b", col="red")
mtext("Cell Density (cells/mL)",side=4,col="red",line=4) 
axis(4, ylim=c(0,7000), col="red",col.axis="red",las=1)

axis(1,pretty(range(filter(data_routine, !is.na(Mean_Nitrite))$Time),10))
mtext("Time (Hours)",side=1,col="black",line=2.5)  

dev.off()
# Export the plot

```

**Section 2 : Nitrososphaera viennensis fast-track cultures activity data**

The plot produced can be found at Supplementary Figure S6A

Step 1: Data input.

```{r Data import and pre-processing, echo=TRUE}
# Here, we create an R data frame named with the data
library(readxl)
data_1D <- read_excel("NVE Standardisation.xlsx", sheet = 2, col_names = TRUE)
# We can change the names of the columns to more R-friendly names
names(data_1D) <- c("Time", "Nitrite", "Mean", "St.Dev", "Repetition")
#We keep only the rows that contain a Nitrite value (thus removing rows that are left blank to make life easier when viewing the .xlsx file)
data_1D <- filter(data_1D, !is.na(Nitrite))
#Let's view the data
glimpse(data_1D)
```

Step 2: Create the plot by overlaying the fast-track culture activity curves with the routine cultures growth curve

```{r Data import and pre-processing, echo=TRUE}
# Create the plot 
line_plot_FTS_vs_routine_activity_NVE <- ggplot(data_1D, aes(x = Time, y = Nitrite)) +
  geom_point(aes(colour = Repetition), size = 2) + 
  geom_smooth(aes(colour = Repetition), method = lm, linetype = "dashed", linewidth = 2, se = F, fullrange = T) + 
  theme_light() + 
  labs(title = "Nitrososphaera viennensis Fast - Track Culture", subtitle = "Linear Regression compared to the growth model for the Routine culture", x = "Time (Hours)", y = "[Nitrite] (μM)") + 
  scale_y_continuous(breaks = seq(from = 0, to = 2400, by = 300), limits = c(0,2400)) +
  scale_x_continuous(breaks = seq(from = 0, to = 250, by = 50), limits = c(0,250)) +
  theme(legend.position = "bottom", legend.box = "vertical", axis.title = element_text(size = 28), axis.text = element_text(size = 24), legend.text = element_text(size = 20), legend.title = element_text(size = 20)) + scale_color_manual(values = c("1st" = "#840084", "2nd" = "#b200b2", "3rd" = "#ff6bff", "Routine Culture" = "#006b6b"), name = "Culture")

line_plot_FTS_vs_routine_activity_NVE <- line_plot_FTS_vs_routine_activity_NVE + 
  geom_point(data = data_routine, aes(x = Time, y = Mean_Nitrite)) + geom_line(data = filter(data_routine, !is.na(Mean_Nitrite)), aes(x = Time, y = Mean_Nitrite, colour = "Routine Culture"), linewidth = 1) 

line_plot_FTS_vs_routine_activity_NVE <- line_plot_FTS_vs_routine_activity_NVE + 
  geom_errorbar(data = filter(data_routine, !is.na(Mean_Nitrite)), aes(x = Time, ymin = Mean_Nitrite - St.Dev_Nitrite, ymax = Mean_Nitrite + St.Dev_Nitrite), colour = "#006b6b", width = 1.5) + 
  geom_text(x = Inf, y = -Inf, label = paste("R - squared =", round(summary(lm(Nitrite~Time, data_1D))$adj.r.squared,4)), colour = "magenta4", size = 10, hjust = 1.1, vjust = -2) 

print(line_plot_FTS_vs_routine_activity_NVE)

cairo_pdf("Figure S6A-N. viennensis activity vs routine.pdf", width = 24, height = 12, onefile = T)
print(line_plot_FTS_vs_routine_activity_NVE)
dev.off()


```

This is the end of this script. 
Please always remember to cite the authors of any code you use!
