---
title: "Root exudates screening"
author: "Alex Kanellopoulos"
date: "2024-12-03"
output: html_document
---

Brief description: This is a script that includes all the necessary code to reproduce the results from the respective manuscript. 

Step 1: Let's see all the necessary packages needed for the analysis and load them. If they are not installed, we will install them first. 

```{r Library Installation and Loading, echo=TRUE, message=FALSE, warning=FALSE}
# Here we store the necessary packages in the necessary_packages variable
necessary_packages <- c("readxl","dplyr","ComplexHeatmap","dendextend","tibble","vegan")

# Here we check if the necessary packages are installed. If not, R installs them 
if (!requireNamespace(necessary_packages, quietly = TRUE)) {
  install.packages(necessary_packages)
}

# Here, for every package included in the necessary_packages variable, R will load it. The character.only = TRUE ensures that library() will try to interpret the package names as characters (as it needs for it to work) 
for (package in necessary_packages) {
  library(package, character.only = TRUE)
}
```

**Section 1 : Plot the heatmap with the AOI% values**

The heatmap is used in Figure 4 

Step 1: Data input.

```{r Data import and pre-processing, echo=TRUE}
library(readxl)
data <- read_xlsx("data.xlsx", sheet = 1)
names(data) <- c("Exudate","N. ureae", "N. multiformis","N. viennensis","Ca. N. franklandianus","Origin")

```

Step 2: Data processing

```{r echo=TRUE, message=FALSE, warning=FALSE}
data_origin <- data.frame(data$Exudate,data$Origin)
names(data_origin) <- c("Exudate","Origin")

library(dplyr)

root_exudates <- data$Exudate
data_AOI <- data[,-c(1,6)]


data_AOI_mat <- as.matrix(data_AOI)
col_to_row_names <- root_exudates
rownames(data_AOI_mat) <- col_to_row_names
```

Step 3: Create the preliminary heatmap

```{r echo=TRUE, message=FALSE, warning=FALSE}

library(ComplexHeatmap)
library(dendextend)

# Create the row and column dendrograms using Euclidean distances. 
row_dend <- as.dendrogram(hclust(dist(data_AOI_mat, method = "euclidean")))
  #Set the row colours number
row_dend <- colour_branches(row_dend, k = 6) 
column_dend <- as.dendrogram(hclust(dist(data_AOI_mat, method = "euclidean"), method = "complete"))
column_dend <- colour_branches(column_dend, k = 1)

RE_AOI_heatmap <- Heatmap(data_AOI_mat, name = "Ammonia Oxidation Inhibition %", 
        column_title = "AOM strains", column_title_side = "bottom", column_title_gp = gpar(fontsize = 16), column_dend_side = "bottom", column_names_gp = gpar(fontsize = 10),
        row_title = "Root Exudate", row_names_side = "left",  row_names_gp = gpar(fontsize = 12),   row_dend_width = unit(10,"mm"), row_title_gp = gpar(fontsize = 16),
        column_title_rot = 0, column_names_rot = 0, column_names_centered = T, 
        col = c("lightblue3", "lightblue1", "yellow", "orange", "orange3", "red"), 
        cluster_rows = row_dend, row_dend_gp = gpar(font = "bold"),
        border = F, 
        column_km = 1, 
        row_split = 6, 
        show_column_dend = T,
        width = ncol(data_AOI_mat)*unit(25,"mm"), height = nrow(data_AOI_mat)*unit(4.5,"mm"), 
        cell_fun = function(j, i, x, y, width, height, fill) {
  grid.text(sprintf("%.2f", data_AOI_mat[i, j]), x, y, gp = gpar(fontsize = 10))
})

# Get the order of the REs as they appear in the clusters from the top to the bottom of the heatmap
row_order_clustered <- row_order(RE_AOI_heatmap)

```

Step 4: Process the data to prepare for annotating the genotype origin

```{r echo=TRUE, message=FALSE, warning=FALSE}

# Collapse the lists of the row order into one list
row_order_clustered_list <- unlist(row_order_clustered, recursive = F)

# Bind the RE order next to the mat used for the heatmap
data_clustered <- cbind(data_AOI_mat, row_order_clustered_list)

# Use the following code to sort the dataframe with the Origin info for all REs
genotype_origin_ordered <- cbind(data_origin, seq(1,20,1))
names(genotype_origin_ordered) <- c("Exudate","Origin","Number")
genotype_origin_ordered <- mutate(genotype_origin_ordered, Number = as.vector(Number))
genotype_origin_ordered$Number <- factor(genotype_origin_ordered$Number, levels = row_order_clustered_list)
genotype_origin_ordered <- genotype_origin_ordered[order(genotype_origin_ordered$Number), ]
View(genotype_origin_ordered)

RE_order <- data.frame(genotype_origin_ordered)$Exudate
genotype_origin_ordered <- genotype_origin_ordered$Origin
origin_order <- genotype_origin_ordered 


library(tibble)
data_AOI_origin <- data_AOI_mat
data_AOI_origin <- data.frame(data_AOI_origin)
data_AOI_origin <- rownames_to_column(data_AOI_origin, var = "Exudate")

data_AOI_origin$Exudate <- factor(data_AOI_origin$Exudate, levels = RE_order)
data_AOI_origin <- data_AOI_origin[order(data_AOI_origin$Exudate), ]
data_AOI_origin <- cbind(seq(1,10,1), data_AOI_origin)
rownames(data_AOI_origin) <- data_AOI_origin$`seq(1, 44, 1)`
rownames(data_AOI_origin) <- data_AOI_origin$Exudate
data_AOI_origin <- data_AOI_origin[,-c(1:2)]

# This is the final mat used for the final heatmap
data_AOI_origin_mat <- as.matrix(data_AOI_origin)

```

Step 5: Prepare the final heatmap

```{r echo=TRUE, message=FALSE, warning=FALSE}
row_ha_Origin = rowAnnotation("Wheat genotype origin" = origin_order, col = list("Wheat genotype origin" = c("India" = "green4", "Austria" = "red")))


row_dend_origin <- as.dendrogram(hclust(dist(data_AOI_origin_mat, method = "euclidean")))
row_dend_origin <- colour_branches(row_dend_origin, k = 10) 

RE_AOI_heatmap_origin_annotated <- Heatmap(data_AOI_origin_mat, name = "Ammonia Oxidation Inhibition %", right_annotation = row_ha_Origin,
                                column_title = "AOM strains", column_title_side = "bottom", column_title_gp = gpar(fontsize = 16), column_dend_side = "bottom", column_names_gp = gpar(fontsize = 10),
                                row_title = "Compound", row_names_side = "left",  row_names_gp = gpar(fontsize = 12), row_dend_width = unit(10,"mm"), row_title_gp = gpar(fontsize = 16),
                                column_title_rot = 0, column_names_rot = 0, column_names_centered = T, 
                                col = c("lightblue3", "lightblue1", "yellow", "orange", "orange3", "red"), 
                                cluster_rows = row_dend_origin, row_dend_gp = gpar(font = "bold"),
                                border = F, 
                                column_km = 1, 
                                row_split = 9, 
                                show_column_dend = T, show_row_dend = T,
                                width = ncol(data_AOI_origin_mat)*unit(25,"mm"), height = nrow(data_AOI_origin_mat)*unit(4.5,"mm"), 
                                cell_fun = function(j, i, x, y, width, height, fill) {
                                  grid.text(sprintf("%.2f", data_AOI_origin_mat[i, j]), x, y, gp = gpar(fontsize = 10))
                                })

RE_AOI_heatmap_origin_annotated

cairo_pdf("Figure 5-Wheat Root Exudates Heatmap.pdf", width = 8, height = 6)
print(RE_AOI_heatmap_origin_annotated)
dev.off()

```

**Section 2 : Plot the PCA with the AOI% values**

The PCA plot is used in Supplementary Figure S12 

Step 1: Data input.

```{r Data import and pre-processing, echo=TRUE}
library(readxl)
data_PCA <- read_xlsx("data.xlsx", sheet = 1)
names(data_PCA) <- c("Exudate","N. ureae", "N. multiformis","N. viennensis","Ca. N. franklandianus","Origin")

```

Step 2: Data processing.

```{r Data import and pre-processing, echo=TRUE}
# Keep only the AOI values
library(dplyr)
data_mat <- data_PCA[, 2:5]

library(vegan)

# Scale the values
data_standardised <- scale(data_mat, center = TRUE, scale = TRUE)

# Insert the exudate numbers as rownames
rownames(data_standardised) <- data_PCA$Exudate

```

Step 3: Perform the PCA.

```{r Data import and pre-processing, echo=TRUE}

library(factoextra)

# Combute the PCA
pca <- prcomp(data_standardised, scale. = TRUE)

# Get the Eigenvalues and store the PCA data
fviz_eig(pca)

pca_data <- as.data.frame(pca$x)
pca_data <- mutate(pca_data, Origin = data_PCA$Origin)

# Plot the PCA bi-plot and print it
REs_biplot <- fviz_pca_biplot(pca,
                              habillage = as.factor(pca_data$Origin),
                              addEllipses = F, ellipse.level = 0.95,
                              title = "PCA bi-plot for Root Exudate Inhibitory pattern",
                              legend.title = "Wheat phenotype origin",
                              repel = T, 
                              palette = c("red", "green4"))
REs_biplot

cairo_pdf("Figure S12-Wheat REs PCA.pdf", width = 8, height = 6, onefile = T)
print(REs_biplot)
dev.off()

```

This is the end of this script. 
Please always remember to cite the authors of any code you use!
